Effective collector efficiency of sparse and dense arrays of emergent cylindrical collectors in laminar-turbulent transitional flows

### 2020-12-14 aa workingdraft/draft.tex

Mechanisms by which plants are known to affect sedimentation include direct capture of sediment on stems, leaves, and other plant surfaces \cite{mudd2010does}, direct deposition of organic matter via root growth and shedding of plant matter (e.g., leaves, seeds, fruit) \cite{nyman2006marsh, neubauer2008contributions}, and effects on turbulence and flow patterns impacting the rate at which suspended particles settle out of the water column due to gravity \cite{christiansen2000flow, leonard1995flow}. Additionally, plants' effects on bedload transport have been the subject of numerous studies \cite[e.g.,][]{yager2013influence, yang2019impact, jordanova2003experimental}. These mechanisms act on different, but sometimes overlapping, portions of a water body's sediment load. For instance, settled particles subsequently become bedload, subject to those vegetative effects. Bedload can also be resuspended by turbulent flow, and vegetation has been found to strongly influence the mechansisms by which this occurs also \cite{tinoco2018turbulence}. Even among suspended particles alone, different mechanisms affect particles with different sizes, specific gravity, and surficial properties. While gravitational settling plays a dominant role for large or dense particles in typical flow conditions for vegetated reaches \cite{mudd2010does, leonard1995flow}, fine particles that travel over great longitudinal distances before settling may be destined to either impact vegetation and be collected directly, or pass through vegetated areas without being retained.

@@

Capture efficiency ($\eta=\frac{b}{d_c}$) is defined as the ratio between the upstream width ($b$) of the streamlines that encounter a collector and the collector's own diameter ($d_c$) (Figure \ref{fig:capeff}). It is a geometrical parameter that describes the ability of a vegetation stem to collect sediment particles from upstream in a given flow field. Effective capture efficiency ($\eta^\prime=p_r\eta$) accounts for the fact that not all particles coming into contact with a collector will adhere to it, thus including probability of retention ($p_r$) as a multiplicand in its calculation. This parameter, as opposed to capture efficiency $\eta$, corresponds to the empirical estimates made by our study and presumably any others that involve $p_r$ as an implicit factor. A formulation that includes efficiency $p_r$ is valuable because it determines the overall ability of a given vegetation system for removing sediment from flow through direct interception.

@@

% Flume diagram figure
\begin{figure}[H]
\includegraphics[width=5in]{../pics/flume_with_sedtraps.png}
\centering
\caption{The Ecogeomorphology flume. Not all measurements are to scale. \textbf{(a)} Photograph of the test section (left-center) and pump (right). \textbf{(b)} Conceptual diagram of the flume as seen from above. Labeled parts are: 1) pump, 2) magnetic flowmeter, 3) test section, and 4) honeycomb flow collimator. Arrows indicate direction of flow. Green points represent the inlets for the peristaltic pumps sampling suspended particle concentration. Red areas represent the settling traps. \textbf{(c)} A side view of the open-channel part of the system.}
\label{fig:floorplan}
\end{figure}

@@

We used 1/8" (0.3175 cm) cylindrical wooden dowels as collectors, which were covered in silicone grease (Chemplex 710, Fuchs Petrolub, Mannheim, Germany) in order to retain impacted particles. Crushed walnut shell was used as a substitute for suspended sediment because it has surficial properties and specific gravity (1.53) comparable to the organic-rich types common in wetlands. We used WF5-200 grade (Composition Materials Co., Milford, CT), which passes entirely through a \#60 sieve (\SI{250}{\micro\metre}), and which we determined has a average particle diameter ($d_{50}$) of \SI{25.2}{\micro\metre} based on measurements using a laser-scattering-based instrument (LISST-Portable|XR, Sequoia Scientific, Bellevue, WA).

@@

% Experimental parameters table
\begin{table}[H]
\caption{Experimental and natural parameter ranges. Natural values are based on approximations or measurements from cited studies and literature reviews.}
\centering
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
\textbf{Parameter}&\textbf{This study}&\textbf{Fauria et al. (2015)}&\textbf{Purich (2006)}&\textbf{Natural}\\
\midrule
Flow velocity (\SI{}{\centi\metre/\second})     
& 2.0--6.0    & 1.8--6.1    & 1.0--10.2    & 0--25 \cite{nikora2008hydraulic}    \\
Flow depth (\SI{}{\centi\metre})                
& 40          & 14--17      & 12           & 0--50 \cite{kadlec1990}    \\
\midrule
Collector shape
& Cylindrical & Artificial grass  & Cylindrical & --- \\ 
Collector diameter (\SI{}{\centi\metre})
& 0.318       & 0.3         & 0.6          & 0.1--1.2 \cite{Nepf_2012,wright2018hydrological} \\
Collector Reynolds number                       
& 66--200     & 54--183     & 70--640      & 5--1000 \cite{kadlec1990}  \\ 
Collector density ($\#/\text{m}^2$)
& 278--1450   & 2724--7209  & 1013--4053   & 10--2700 \cite{wright2018hydrological} \\
Solid volume fraction (\%)
& 0.22--1.17  & 0.82--2.16  & 2.86--11.5   & 0.1--1 \cite{Nepf_2012}   \\ 
\midrule
Particle type
& Walnut shell  & Road dust  & Pliolite\textsuperscript{\textregistered}   & --- \\ 
Specific gravity      
& 1.53        &2.27--2.61 \tnote{1}  & 1.03         & 2.56--2.87 \cite{czuba2015sedgravity} \\
Average particle diameter, $d_{50}$ (\SI{}{\micro\metre})     
& 25.2        & $\sim$10--15      & 212--250     & 45--100 \cite{hejduk2010variations,noe2010glades}  \\
Suspended concentration (\SI{}{\micro\liter/\liter})      
& 5.3--53.8   & 9--50       & $\sim$110   & 2-25 \cite{noe2010glades,aiona2013can}      \\
\midrule
Particle-collector diameter ratio, $R$      
&0.0079       &0.0004--0.083 \tnote{2} & 0.037        & <0.25     \\
\bottomrule
\vspace{-4mm}
\end{tabular}
\begin{tablenotes}
\footnotesize \item[1] Estimated from a different study about road dust properties \cite{mckenzie2008size} 
\vspace{2mm}
\footnotesize \item[2] Fauria et al. \cite{Fauria_2015} used a LISST to measure collection across a broad range of particle diameters (1.25--250 \SI{}{\micro\metre})
\end{tablenotes}
\end{threeparttable}
\label{tbl:parameters}
\end{table}

@@

We estimated mid-water-column turbulent kinetic energy (TKE) among the collectors for each combination of $\Rey_c$ and collector density using the Vectrino Profiler ADV. The Vectrino was mounted on a frame suspended above the test section and adjusted until its bottom-depth measurement read 15 cm. A bubble level was used to ensure it was oriented along the vertical axis. The ADV measured velocity in two horizontal dimensions, and two separate measurements along the vertical dimension, at a frequency of 10 Hz. It also collected other data such as autocorrelation and signal-to-noise ratio. We collected 5 minutes of data per treatment to ensure we captured any possible periodicity in turbulence. Measurements taken by the ADV were filtered, leaving out timepoints where the two vertical velocity measurements differed by 50\% or more or the SNR measurement was greater than 5 dB, and were subsequently despiked using a phase-space thresholding algorithm \cite{goring_nikora}. TKE was derived from the cleaned Vectrino data as half the sum of the variances of the velocity components:
\begin{equation}
    \text{TKE} = \frac{1}{2}[\overline{(u^\prime)^2} + \overline{(v^\prime)^2} + \overline{(w^\prime)^2}]\,.
    \label{eqn:TKE}
\end{equation}

@@

The effect of biofilm on the surficial properties of collectors, and thereby their effective collector efficiency, was another factor of interest for our experiment. While we did not incorporate biofilm into the fully-crossed parameter space of other explicative variables, we carried out an auxiliary experiment to assess its effect size relative to the other variables. After one of the experiments at each collector density, the water in the flume was allowed to come to rest, then the collectors sat constantly illuminated by one white, fluorescent overhead light for a number of days. The crushed walnut shell doubled as a food source for the microorganisms constituting the biofilm, which we observed grew plentifully in all cases.

Once biofilm became visually apparent on the collectors' surfaces, an experiment was conducted using the same materials and protocols as the treatments without biofilm, holding $\Rey_c$ constant between biofilm treatments (200). The length of time it took for robust biofilm growth to occur differed between the low, medium, and high collector-density treatments (13; 18; 20 days). For the highest collector-density treatment, the flume was left full and illuminated for a longer period (46 days), after which a second experiment with more biofilm present was performed. While we did not quantitatively assess biofilm mass, we assumed that the number of days it was allowed to grow could be used as a proxy, acknowledging that the growth rate (in mass units per day) was probably exponential or logistical rather than linear.

@@

\noindent $\bar{\phi}(t) = \frac{1}{h} \int_0^h\phi(z)dz$ is the depth-averaged suspended sediment concentration at time $t$, where $h$ is water depth and $z$ is the vertical dimension. $k_s = \frac{Cv_s}{h}(1-E_r)$ is the exponential time-decay rate of $\bar{\phi}$ due to settling, which is calculated from settling velocity ($v_s$), water depth ($h$), a constant $C$ that accounts for the shape of the Rouse profile, and a dimensionless rate at which particles are entrained from the bed ($E_r$). Because we focused on particle capture, we only empirically estimated $k_s$, not its constituent terms. However, further derivation is available in Fauria et al. \cite{Fauria_2015}. $k_c$ = $\eta^{\prime}ud_cI_c$ is the rate at which particles are removed from suspension due to capture by collectors. It is the product of effective capture efficiency, flow velocity ($u$), collector diameter ($d_c$), and collector height density ($I_c$). For emergent collectors, $I_c = \frac{N_ch}{V}$, where $N_c$ is the total number of stems, and $V$ is the volume of water in which stems are contained. The product $d_cI_c$ is also referred to as frontal-area density ($a$).

To derive $k_c$, begin by considering a single collector. Its rate of collection is the product of the probability of retention ($p_r$) and the flux of particles passing through the cross-section ($b \times h$) where streamlines reach its surface, $p_r ubh\bar{\phi} = p_r \eta u d_c h \bar{\phi}$. By substituting for effective collector efficiency ($\eta^\prime = p_r \eta$) and multiplying by the number of collectors acting on the suspended particles per unit volume, the $k_c$ term is derived: $\eta^\prime u d_c \frac{N_c h}{V} \bar{\phi} = \eta^{\prime}ud_cI_c \bar{\phi} = k_c \bar{\phi}$.

@@

\noindent Because we assume the vertical profile of sediment concentration maintains a constant shape through time, governed by the Rouse equation, concentration at any height ($\phi_a$) can be assumed to be proportional to average concentration ($\phi_a=\frac{\bar{\phi}}{C}$):
\begin{equation}
    \phi_a(t) = \phi_a(0)e^{-kt}\,,
    \label{eq:specconc}
\end{equation}
which is the formula we used to estimate total time-decay of suspended concentration from our point measurements in the vertical profile.

To calculate $k_s$, we began by calculating the average mass of sediment that settled per settling trap over the duration of the experiment (100 minutes). We then scaled this measurement up from the area of a settling trap's horizontal opening (5.07 cm$^2$) to the total test section bed area (1.17 m$^2$). The resulting estimate of the sediment mass settled in the test section ($m_s$) was divided by the total mass removed from suspension during a run to calculate the proportion of the decrease due to settling. This proportion was multiplied by the total exponential decay coefficient, $k$, to calculate $k_s$:

\begin{equation}
    k_s = \frac{m_s}{m_0(1-e^{-kT})}k\,,
\end{equation}

@@

The equations above model the dynamics occurring in the test section alone. To account for settling and re-entrainment outside the test section, a background exponential decay rate ($k_b$) was estimated from runs with zero collectors for each $\Rey_c$ treatment ($k_c = 0$, hence $k = k_s + k_b$). ECE for our collector-density treatments was then calculated from $k_c$, which was calculated from $k$, $k_s$, and $k_b$:

\begin{equation}
    \eta^\prime = \frac{k_c}{ud_cI_c} = \frac{k - k_s - k_b}{ud_cI_c}\,.
\end{equation}

@@

\subsubsection{Contextualizing Findings}

@@

Effective collector efficiency was found to decrease with increasing collector Reynolds number and increasing collector density throughout our parameter space, with one exception. This exception occurred at the lowest collector density ($\phi_c = 0.22\%$), where the intermediate $\Rey_c$ treatment ($\Rey_c = 134$) had lower ECE than both the lower and higher $\Rey_c$ treatments (Figure \ref{fig:ece}a). When compared to the other collector-density treatments with the same $\Rey_c$ value, this low collector-density experiment's calculated ECE was slightly lower than the intermediate collector-density treatment, and was greater than the highest collector-density treatment (Figure \ref{fig:ece}b). For both of these comparisons, the large magnitude of our uncertainty leaves open the possibility that the mid-$\Rey_c$, low-collector-density result could simply be an anomaly.

@@

Monte Carlo analysis determined that there is a significant (null hypothesis: $\mu = 0$) negative relationship between ECE and collector density for the low (p = 0.004) and high (p = 0.019) $\Rey_c$ treatments, but no significant effect was detected (p > 0.05) for the intermediate treatment (Figure \ref{fig:monte}a). We also found a significant negative relationship between ECE and $\Rey_c$ for the low (p = 0.026) and high (p = 0.046) collector-density treatments, but an insignificant (p > 0.05) trend for the intermediate collector-density treatment (Figure \ref{fig:monte}b).

% Monte Carlo figure
\begin{figure}[H]
\centering
\includegraphics[width=5in]{../pics/montecarlo.png}
\caption{Effect size estimates resulting from Monte Carlo regression analysis. Error bars represent 95\% confidence intervals. Asterisks mark the parameter values at which the effect size was found to be significant (\textbf{a}) Estimates of the effect of log-transformed collector solid volume fraction ($\beta_{\log(\phi_c)}$) on ECE, stratified by $\Rey_c$. (\textbf{b}) Estimates of the effect of log-transformed $\Rey_c$ ($\beta_{\log(\Rey_c)}$) on ECE, stratified by solid volume fraction ($\phi_c$).}
\label{fig:monte}
\end{figure}

@@

Turbulent kinetic energy was consistently found to increase with increasing flow velocity as expected (Table \ref{tbl:turbulence}). Generally, higher collector density also led to greater turbulence, with the greatest increase in terms of both absolute and relative magnitude taking place between the lower and middle collector-density treatments. A slight decrease actually occurred thereafter (between the middle and highest collector densities) for the middle and highest $\Rey_c$ treatments, with a slight increase for the lowest $\Rey_c$ treatment. Altogether, these results likely indicate that the turbulent eddy scale and the distance between collectors converged somewhere in our range for those parameters, so that additional collectors acted to damp eddies more than they acted to create them. If there is a threshold for this effect, our results, including the similarity between the lowest collector density and the zero-collector control, suggest that it would occur at a collector-density value between our lower and middle treatments.

% Turbulence table
\begin{table}[H]
\caption{Turbulence and flow characteristics for all collector solid volume fraction ($\phi_c$) and $\Rey_c$ treatments, including the control (zero-collector) treatments for reference.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{>{\bfseries}r>{\bfseries}rrr}
\toprule
\textbf{$\phi_c$}&\textbf{$\Rey_c$}&\textbf{mid-water-column TKE (\SI{}{\milli\metre^2/\second^2})}\\
\midrule
Control &   67  & \num{2.48}\\
        &   134 & \num{2.98}\\
        &   200 & \num{3.37}\\
\midrule
0.22\% &   67  & \num{1.23}\\
        &   134 & \num{5.26}\\
        &   200 & \num{12.2}\\
\midrule
0.64\% &   67  & \num{9.13}\\
        &   134 & \num{35.9}\\
        &   200 & \num{59.6}\\
\midrule
1.17\%  &   67  & \num{11.1}\\
        &   134 & \num{29.8}\\
        &   200 & \num{54.4}\\
\bottomrule
\label{tbl:turbulence}
\end{tabular}
\end{table}

@@

% Fauria model comparison figure
\begin{figure}[H]
\centering
\includegraphics[width=5in]{../pics/comparisonplot.png}
\caption{A comparison of our experimental estimates of effective capture efficiency, colored according to collector solid volume fraction ($\phi_c$), and the predictions of the Fauria et al. \cite{Fauria_2015} power law model. Effective capture efficiency is plotted on a log scale in order to display model fit more accurately at small values. Error bars represent +/- 1 SE.}
\label{fig:compplot}
\end{figure}

>>>>

Mechanisms by which plants are known to affect sedimentation include direct capture of sediment on stems, leaves, and other plant surfaces \cite{mudd2010does}, direct deposition of organic matter via root growth and shedding of plant matter (e.g., leaves, seeds, fruit) \cite{nyman2006marsh, neubauer2008contributions}, and effects on turbulence and flow patterns impacting the rate at which suspended particles settle out of the water column due to gravity \cite{christiansen2000flow, leonard1995flow}. Additionally, plants' effects on bedload transport have been the subject of numerous studies \cite[e.g.,][]{yager2013influence, yang2019impact, jordanova2003experimental}. These mechanisms act on different, but sometimes overlapping, portions of a water body's sediment load. For instance, settled particles subsequently become bedload, subject to those vegetative effects. Bedload can also be resuspended by turbulent flow, and vegetation has been found to strongly influence the mechanisms by which this occurs also \cite{tinoco2018turbulence}. Even among suspended particles alone, different mechanisms affect particles with different sizes, specific gravity, and surficial properties. While gravitational settling plays a dominant role for large or dense particles in typical flow conditions for vegetated reaches \cite{mudd2010does, leonard1995flow}, fine particles that travel over great longitudinal distances before settling may be destined to either impact vegetation and be collected directly, or pass through vegetated areas without being retained.

@@

Capture efficiency ($\eta=\frac{b}{d_c}$) is defined as the ratio between the upstream width ($b$) of the streamlines that encounter a collector and the collector's own diameter ($d_c$) (Figure \ref{fig:capeff}). It is a geometrical parameter that describes the ability of a vegetation stem to collect sediment particles from upstream in a given flow field. Effective capture efficiency ($\eta^\prime=p_r\eta$; ECE) accounts for the fact that not all particles coming into contact with a collector will adhere to it, thus including probability of retention ($p_r$) as a multiplicand in its calculation. ECE is the appropriate variable for the empirical estimates made by our study and presumably any others that involve $p_r$ as an implicit factor, and it, rather than $\eta$, corresponds to the realized efficiency of natural vegetation for removing sediment from flow through direct interception.

@@

% Flume diagram figure
\begin{figure}[H]
\includegraphics[width=5in]{../pics/flume_with_sedtraps.png}
\centering
\caption{The Ecogeomorphology flume. Not all measurements are to scale. \textbf{(a)} Photograph of the test section (center-left) and pump (right). \textbf{(b)} Conceptual diagram of the flume as seen from above. Labeled parts are: 1) pump, 2) magnetic flowmeter, 3) test section, and 4) honeycomb flow collimator. Arrows indicate direction of flow. Green points represent the inlets for the peristaltic pumps sampling suspended particle concentration. Red areas represent the settling traps. \textbf{(c)} A side view of the open-channel part of the system.}
\label{fig:floorplan}
\end{figure}

@@

We used 1/8" (0.3175 cm) cylindrical wooden dowels as collectors, which were covered in silicone grease (Chemplex 710, Fuchs Petrolub, Mannheim, Germany) in order to retain impacted particles. Crushed walnut shell was used as a substitute for suspended sediment because it has surficial and physical properties comparable to the organic-rich types common in wetlands \cite{muller2017experiments, jenzer2015sediment, redding2006particle}. We used WF5-200 grade (Composition Materials Co., Milford, CT), which passes entirely through a \#60 sieve (\SI{250}{\micro\metre}), and which we determined has a average particle diameter ($d_{50}$) of \SI{25.2}{\micro\metre} based on measurements using a laser-scattering-based instrument (LISST-Portable|XR, Sequoia Scientific, Bellevue, WA). We also empirically estimated the particle density for this specific walnut shell flour to be 1.53 \SI{}{\gram/\centi\metre\cubed}.

@@

% Experimental parameters table
\begin{table}[H]
\caption{Experimental and natural parameter ranges. Natural values are based on approximations or measurements from cited studies and literature reviews in wetlands or other low-elevation flows.}
\centering
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
\textbf{Parameter}&\textbf{This study}&\textbf{Fauria et al. (2015)}&\textbf{Purich (2006)}&\textbf{Natural}\\
\midrule
Flow velocity (\SI{}{\centi\metre/\second})     
& 2.0--6.0    & 1.8--6.1    & 1.0--10.2    & 0--25 \cite{nikora2008hydraulic}    \\
Flow depth (\SI{}{\centi\metre})                
& 40          & 14--17      & 12           & 0--50 \cite{kadlec1990}    \\
\midrule
Collector shape
& Cylindrical & Artificial grass  & Cylindrical & Varies \\ 
Collector diameter (\SI{}{\centi\metre})
& 0.318       & 0.3         & 0.6          & 0.1--1.2 \cite{Nepf_2012,wright2018hydrological} \\
Collector Reynolds number                       
& 66--200     & 54--183     & 70--640      & 5--1000 \cite{kadlec1990}  \\ 
Collector density ($\#/\text{m}^2$)
& 278--1450   & 2724--7209  & 1013--4053   & 10--2700 \cite{wright2018hydrological} \\
Solid volume fraction (\%)
& 0.22--1.17  & 0.82--2.16  & 2.86--11.5   & 0.1--1 \cite{Nepf_2012}   \\ 
\midrule
Particle type
& Walnut shell  & Road dust  & Pliolite\textsuperscript{\textregistered}   & Sediment \\ 
Particle density  (\SI{}{\gram/\centi\metre\cubed})    
& 1.53        &2.27--2.61 \tnote{1}  & 1.03         & 1.43--2.39 \cite{redding2006particle} \\
Average particle diameter, $d_{50}$ (\SI{}{\micro\metre})     
& 25.2        & $\sim$10--15 \tnote{2} & 212--250     & 45--100 \cite{hejduk2010variations,noe2010glades}  \\
Suspended concentration (\SI{}{\micro\liter/\liter})      
& 5--55   & 9--50       & $\sim$110   & 2-25 \cite{noe2010glades,aiona2013can}      \\
\midrule
Particle-collector diameter ratio, $R$      
&0.0079       &0.0004--0.083 \tnote{2} & 0.037        & <0.25     \\
\bottomrule
\vspace{-4mm}
\end{tabular}
\begin{tablenotes}
\footnotesize \item[1] Estimated from a different study about road dust properties \cite{mckenzie2008size} 
\vspace{2mm}
\footnotesize \item[2] Fauria et al. \cite{Fauria_2015} used a LISST to measure collection across a broad range of particle diameters (1.25--250 \SI{}{\micro\metre})
\end{tablenotes}
\end{threeparttable}
\label{tbl:parameters}
\end{table}

@@

We estimated mid-water-column turbulent kinetic energy (TKE) among the collectors for each combination of $\Rey_c$ and collector density using the Vectrino Profiler ADV. The Vectrino was mounted on a frame suspended above the test section and adjusted until its bottom-depth measurement read 15 cm. A bubble level was used to ensure it was oriented along the vertical axis. The ADV measured velocity in two horizontal dimensions, and two separate measurements along the vertical dimension, at a frequency of 10 Hz. It also collected other data such as autocorrelation and signal-to-noise ratio. We collected 5 minutes of data per treatment in the interest of including the vast majority of expected periodicity in our variance measurement. Measurements taken by the ADV were filtered, leaving out timepoints where the two vertical velocity measurements differed by 50\% or more or the SNR measurement was greater than 5 dB, and were subsequently despiked using a phase-space thresholding algorithm \cite{goring_nikora}. TKE was derived from the cleaned Vectrino data as half the sum of the variances of the velocity components:
\begin{equation}
    \text{TKE} = \frac{1}{2}[\overline{(u^\prime)^2} + \overline{(v^\prime)^2} + \overline{(w^\prime)^2}]\,.
    \label{eqn:TKE}
\end{equation}

@@

The effect of biofilm on the surficial properties of collectors, and thereby their effective capture efficiency, was another factor of interest for our experiment. While we did not incorporate biofilm into the fully-crossed parameter space of other explicative variables, we carried out an auxiliary experiment to roughly assess its effect size. After one of the experiments at each collector density, the water in the flume was allowed to come to rest, then the collectors sat constantly illuminated by one white, fluorescent overhead light for a measured number of days. The crushed walnut shell doubled as a food source for the microorganisms constituting the biofilm, which we observed grew plentifully in all cases.

Once biofilm became visually apparent on the collectors' surfaces, an experiment was conducted using the same materials and protocols as the treatments without biofilm, holding $\Rey_c$ constant between biofilm treatments (200). The length of time it took for robust biofilm growth to occur differed between the low, medium, and high collector-density treatments (13; 18; 20 days). For the highest collector-density treatment, the flume was left full and illuminated for a longer period (46 days), after which a second experiment with more biofilm present was performed. While we did not quantitatively assess biofilm mass, we assumed that the number of days it was allowed to grow could be used as a proxy, acknowledging that the growth rate (in mass units per day) was probably exponential or logistic rather than linear.

@@

\noindent $\bar{\phi}(t) = \frac{1}{h} \int_0^h\phi(z)dz$ is the depth-averaged suspended sediment concentration at time $t$, where $h$ is water depth and $z$ is a spatial coordinate along the vertical axis. $k_s = \frac{Cv_s}{h}(1-E_r)$ is the exponential time-decay rate of $\bar{\phi}$ due to settling, which is calculated from settling velocity ($v_s$), water depth ($h$), a constant $C$ that accounts for the shape of the Rouse profile, and a dimensionless rate at which particles are entrained from the bed ($E_r$). Because we focused on particle capture, we only empirically estimated $k_s$, not its constituent terms. However, further derivation is available in Fauria et al. \cite{Fauria_2015}. $k_c$ = $\eta^{\prime}ud_cI_c$ is the rate at which particles are removed from suspension due to capture by collectors. It is the product of effective capture efficiency, flow velocity ($u$), collector diameter ($d_c$), and collector height density ($I_c$). For emergent collectors, $I_c = \frac{N_ch}{V}$, where $N_c$ is the total number of stems, and $V$ is the volume of water in which stems are contained. The product $d_cI_c$ is also referred to as frontal-area density ($a$).

To derive $k_c$, begin by considering a single collector. Its rate of collection is the product of the probability of retention ($p_r$) and the flux of particles passing through the cross-section ($b \times h$) where streamlines reach its surface, $p_r ubh\bar{\phi} = p_r \eta u d_c h \bar{\phi}$. By substituting for effective capture efficiency ($\eta^\prime = p_r \eta$) and multiplying by the number of collectors acting on the suspended particles per unit volume, the $k_c$ term is derived: $\eta^\prime u d_c \frac{N_c h}{V} \bar{\phi} = \eta^{\prime}ud_cI_c \bar{\phi} = k_c \bar{\phi}$.

@@

Concentration at any height ($\phi_a$) can be assumed to change in proportion to average concentration because we assume the vertical profile of sediment concentration maintains a constant shape through time, governed by the Rouse equation \cite{rouse1937modern},

\begin{equation}
    \frac{\phi}{\phi_a} = \left( \frac{h - z}{z} \frac{a}{h - a} \right)^{R_0}\,,
    \label{eq:rouse}    
\end{equation}

\noindent where $\phi$ is concentration at height $z$, and $\phi_a$ is concentration at an arbitrary reference height $a$, and $R_0$ is referred to as the Rouse number \cite{kumbhakar2017derivation}.

To calculate $k_s$, we began by calculating the average mass of sediment that settled per settling trap over the duration of the experiment (100 minutes). We then scaled this measurement up from the area of a settling trap's horizontal opening (5.07 $\times 10^{-4}$ m$^2$) to the total test section bed area (1.17 m$^2$). The resulting estimate of the sediment mass settled in the test section ($m_s$) was divided by the total mass removed from suspension during a run to calculate the proportion of the decrease due to settling. This proportion was multiplied by the total exponential decay coefficient, $k$, to calculate $k_s$:

\begin{equation}
    k_s = \frac{m_s}{m_0(1-e^{-kT})}k\,,
    \label{eq:ks}
\end{equation}

@@

The equations above model the dynamics occurring in the test section alone. To account for settling and re-entrainment outside the test section, a background exponential decay rate ($k_b$) was estimated from runs with zero collectors for each $\Rey_c$ treatment ($k_c = 0$, hence $k = k_s + k_b$). ECE for our collector-density treatments was then calculated from $k_c$, which was calculated from $k$, $k_s$, and $k_b$. Also, because collectors only act on suspended particles while they are passing through the test section, this estimate for ECE was scaled by the ratio (5.19) between the total volume of water in the flume (\SI{2.43}{\metre\cubed}) and the volume of water in the test section (\SI{0.468}{\metre\cubed}):

\begin{equation}
    \eta^\prime = 5.19\frac{k_c}{ud_cI_c} = 5.19\frac{k - k_s - k_b}{ud_cI_c}\,.
    \label{eq:eta}
\end{equation}

\noindent Using this volumetric scaling factor to estimate actual ECE, as Fauria et al. also did \cite{Fauria_2015}, is justified under the assumption that evapotranspiration and any other fluxes in and out of the flume's water volume are negligible.

@@

In order to correct for observed heteroscedasticity across time, with earlier, higher-concentration measurements showing greater variance, we log-transformed the exponential model (Eq. (\ref{eq:expo})). This resulted in a linear model taking the form,

\begin{equation}
    log(\boldsymbol{\phi_a}) = log(\phi_a(0)) - k\boldsymbol{t} + \boldsymbol{U} \,,
\end{equation}

\noindent which was evaluated for each experimental configuration separately. Bold font signifies vector variables, and $\boldsymbol{U}$ is the random error term of the model. We used residual plots to verify that this transformed model exhibited uniformity of residuals (See Supplementary Information). The standard error of this $k$ term and the standard deviation of $m_s$ estimates from settling traps were propagated through Eqs. (\ref{eq:ks}) and (\ref{eq:eta}), for both control and treatment runs, and were ultimately used to calculate uncertainty estimates for ECE.

Our experiment collected suspended concentration samples at three different heights and both upstream and downstream of the test section, whereas Fauria et al. \cite{Fauria_2015} collected at one position only. Because the observed magnitude of the Rouse profile (Eq. (\ref{eq:rouse})) and the upstream-downstream difference were barely detectable with our degree of precision, and both were too small to substantially affect model results, we opted to pool data from all six suspended sediment sampling locations rather than complicating the model further.

To assess the significance of the effects on ECE of $\Rey_c$ and collector density, the latter parameterized as solid volume fraction ($\phi_c$) in the interest of comparability with other studies, we used a Monte Carlo approach. For each combination of $\Rey_c$ and $\phi_c$, 30,000 random values were drawn from a normal distribution whose mean equals that combination's empirically estimated ECE and whose standard deviation equals its standard error (SE). These values were log-transformed, since we expected the relationships to fit a power-law model. For each stratum of a given explicative variable, sets (triplets) with one randomly sampled value from each of the strata of the other explicative variable were regressed using ordinary least squares, producing a roughly normal distribution of coefficients (n = 30,000) for each level of both variables (6 total). The distributions of slopes were used in two-tailed t-tests (null hypothesis: $\mu = 0$; $\alpha = 0.05$) to determine the significance of relationships.

@@

Effective capture efficiency was found to decrease with increasing collector Reynolds number and increasing collector density throughout our parameter space, with one exception. This exception occurred at the lowest collector density ($\phi_c = 0.22\%$), where the intermediate $\Rey_c$ treatment ($\Rey_c = 134$) had lower ECE than both the lower and higher $\Rey_c$ treatments (Figure \ref{fig:ece}a). When compared to the other collector-density treatments with the same $\Rey_c$ value, this low collector-density experiment's calculated ECE was slightly lower than the intermediate collector-density treatment, and was greater than the highest collector-density treatment (Figure \ref{fig:ece}b). Our experimental uncertainty was great enough that this might be attributable to random error. Surprisingly, repeating the experiment with the same parameters yielded a nearly identical result (omitted from analysis).

@@

Monte Carlo analysis determined that there is a significant negative relationship between ECE and collector density for the low (p = 0.004) and high (p = 0.019) $\Rey_c$ treatments, but no significant effect was detected (p > 0.05) for the intermediate treatment (Figure \ref{fig:monte}a). We also found a significant negative relationship between ECE and $\Rey_c$ for the low (p = 0.026) and high (p = 0.046) collector-density treatments, but an insignificant (p > 0.05) trend for the intermediate collector-density treatment (Figure \ref{fig:monte}b).

% Monte Carlo figure
\begin{figure}[H]
\centering
\includegraphics[width=5in]{../pics/montecarlo.png}
\caption{Effect size estimates from Monte Carlo regression analysis. Error bars represent 95\% confidence intervals. Asterisks mark the parameter values at which the effect size was found to be significant. (\textbf{a}) Estimates of the effect of log-transformed collector solid volume fraction ($\beta_{\log(\phi_c)}$) on ECE, stratified by $\Rey_c$. (\textbf{b}) Estimates of the effect of log-transformed $\Rey_c$ ($\beta_{\log(\Rey_c)}$) on ECE, stratified by solid volume fraction ($\phi_c$).}
\label{fig:monte}
\end{figure}

@@

Turbulent kinetic energy was consistently found to increase with increasing flow velocity as expected (Table \ref{tbl:turbulence}). Generally, higher collector density also led to greater turbulence, with the greatest increase in terms of both absolute and relative magnitude taking place between the lower and middle collector-density treatments. A slight decrease actually occurred thereafter (between the middle and highest collector densities) for the middle and highest $\Rey_c$ treatments, with a slight increase for the lowest $\Rey_c$ treatment. Altogether, these results likely indicate that the turbulent eddy scale and the distance between collectors converged somewhere in our range for those parameters, so that additional collectors acted to damp eddies more than they acted to create them. If there is a threshold for this effect, our results, including the relative similarity between the lowest collector density and the zero-collector control, suggest that it would occur at a collector-density value between our lower and middle treatments.

% Turbulence table
\begin{table}[H]
\caption{Turbulent kinetic energy (TKE) for all collector solid volume fraction ($\phi_c$) and $\Rey_c$ treatments, including the control (zero-collector) treatments for reference.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{>{\bfseries}r>{\bfseries}rrr}
\toprule
\textbf{$\phi_c$}&\textbf{$\Rey_c$}&\textbf{mid-water-column TKE (\SI{}{\milli\metre^2/\second^2})}\\
\midrule
Control &   67  & \num{2.48}\\
        &   134 & \num{2.98}\\
        &   200 & \num{3.37}\\
\midrule
0.22\% &   67  & \num{1.23}\\
        &   134 & \num{5.26}\\
        &   200 & \num{12.2}\\
\midrule
0.64\% &   67  & \num{9.13}\\
        &   134 & \num{35.9}\\
        &   200 & \num{59.6}\\
\midrule
1.17\%  &   67  & \num{11.1}\\
        &   134 & \num{29.8}\\
        &   200 & \num{54.4}\\
\bottomrule
\label{tbl:turbulence}
\end{tabular}
\end{table}

@@

% Fauria model comparison figure
\begin{figure}[H]
\centering
\includegraphics[width=5in]{../pics/comparisonplot.png}
\caption{A comparison of our experimental estimates of effective capture efficiency, colored according to collector solid volume fraction ($\phi_c$), and the predictions of the Fauria et al. \cite{Fauria_2015} power law model (black lines). Effective capture efficiency is plotted on a log scale in order to display model fit more accurately at small values. Error bars represent +/- 1 SE.}
\label{fig:compplot}
\end{figure}

Having confirmed that the Fauria et al. \cite{Fauria_2015} model is a reasonably good predictor for our results, we found the $C$ values for our different collector-density treatments and compared them to those estimated for the two collector densities studied by Fauria et al. ($\phi_c$ = 0.82\%, 2.16\%). We then fit a preliminary power-law regression to the results at the two studies' five combined $\phi_c$ test values, which fit relatively well (R$^2$ = 0.82) though data scarcity and the differing environments of the two studies are important to consider in evaluating the certainty of this model's coefficients.
